---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# visualjoin

<!-- badges: start -->
<!-- badges: end -->

The goal of visualjoin is to ...

Sometimes joins feel awkward because you need to know what variables are in the constituent data frames in order to coordinate a good join.  But if you are in a pipeline already, then these dataframes aren't printed for you.  Below, we just get an error - oh no! (unless there is a natural join that can happen - common variables)

```{r cars, error = T}
library(tidyverse)

unvotes::un_votes |>
  left_join(y = rnaturalearth::countries110)
```


Vizjoin will, allow you to *not* have everything declared at once, but will show you helpful dataframe summaries as you define your join




```{r, eval = F}
unvotes::un_votes |>
  viz_left_join()
# returns x data frame variables (glimpse-like)


unvotes::un_votes |>
  viz_left_join(y = rnaturalearth::countries110)
# returns x and y data frame variables (glimpse-like) if no natural join happens
```




---


# first we want to have a function that is kind of like glimpse, only returns a tibble...


```{r}
data_to_df_summary <- function(data = unvotes::un_votes,
  num_values = 4,
  num_columns = ncol(data),
  max_char = 10){
  
data |>
  slice(1:num_values) |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(col = 1:num_columns) |>
  group_by(name) |>
  summarise(values = paste(value, collapse = ", ")) |>
  # mutate(values = str_extract(values))
  mutate(row_number = row_number()) |>
  select(row_number, everything())
  
}

data_to_df_summary()
```


# Then, slightly modifying, we'll have it return different column names depending on if it is right or left side dataframe.

```{r}
x_data_to_df_summary <- function(data = unvotes::un_votes,
                                 num_values = 4){
  
  data |> 
    data_to_df_summary() |>
    rename(x_vars = name,
           x_values = values) 
  
  
}

x_data_to_df_summary()
```


```{r}
y_data_to_df_summary <- function(
    data = unvotes::un_votes,
    num_values = 4){
  
  data |> 
    data_to_df_summary() |>
    rename(y_vars = name,
           y_values = values) 
  
  
}


y_data_to_df_summary()

rnaturalearth::countries110 |>
  sf::st_drop_geometry() |>
  y_data_to_df_summary()
```


```{r}

viz_left_join <- function(x, y = NULL, by = NULL, natural_join = F, ...){
  
  if(is_null(y)){out <- x_data_to_df_summary(x)}
  
  if(is_null(by)&!is.null(y)&!natural_join){
    
    out <- full_join(x_data_to_df_summary(x),
              y_data_to_df_summary(y)) |>
      select(-row_number)
   
    message("add 'by' in this form:  `by = join_by(x_var_name == y_var_name)`")
     
  }
  
  if(!is_null(by)&!is.null(y)){
  
  out <- left_join(x, y, by, ...)
    
  }
  
  if(natural_join){
    
  out <- left_join(x, y, ...)
    
  }
  
  out
  
}
```

#  And now the traditional readme

## You can look at your x data summary

```{r}
unvotes::un_votes %>% 
  viz_left_join()
```

## ... and then along y data summary

```{r}
unvotes::un_votes %>% 
  viz_left_join(y = gapminder::gapminder)
```


## and performing the join (not very interesting - need a better example!)

```{r}
unvotes::un_votes %>% 
  viz_left_join(y = gapminder::gapminder,
                by = join_by(country == country))
```

## And left_join() arguments are usable.

```{r}
unvotes::un_votes %>% 
  viz_left_join(y = gapminder::gapminder,
                by = join_by(country == country), 
                relationship = "many-to-many")

```


# And here we do a bit smaller join...

```{r}
unvotes::un_votes %>% 
  filter(rcid == 3) %>% 
  viz_left_join(y = gapminder::gapminder %>% 
                  filter(year == min(year)),
                by = join_by(country == country))
```

# And we can specify that a natural join is allowed.

```{r}
unvotes::un_votes |>
  viz_left_join(unvotes::un_roll_calls,
                natural_join = T)

```

# And join sequences


```{r}
unvotes::un_votes |>
  viz_left_join(unvotes::un_roll_calls)

unvotes::un_votes |>
  viz_left_join(unvotes::un_roll_calls,
                natural_join = T)


unvotes::un_votes |>
  viz_left_join(unvotes::un_roll_calls,
                natural_join = T) %>% 
  mutate(year = year(date)) %>% 
  viz_left_join(gapminder::gapminder)


unvotes::un_votes |>
  viz_left_join(unvotes::un_roll_calls,
                natural_join = T) %>% 
  mutate(year = year(date)) %>% 
  viz_left_join(gapminder::gapminder, 
                natural_join = T)


```

